### 2024-04-01 Reply-Reviewer2.tex
reply to reviewer 2 complete

%  -----------------------------------------------------------------------------
%  Modified: 17 December 2023
%  -----------------------------------------------------------------------------
\documentclass{letter}

 \date{\today}

\usepackage{color}

\begin{document}
\begin{letter}{
Physical Review Fluids\\
Professor Eric S. G. Shaqfeh, Editor\\}

\opening{Dear Prof. Shaqfeh,}

Please find enclosed the revised version of our manuscript (FZ10233):
``Numerical Simulation of Bubble Deformation and Breakup under 
Simple Linear Shear Flows''
We thank the reviewers for their comments and thoughts regarding improvement 
of our paper. We believe that we have addressed all of the reviewers’ 
concerns; the changes are itemized in detail below.

\par\noindent
NOTE: all references to equations and figures below are with
reference to the numbering scheme of the revised version of the paper,
not the original version.  Also, changes are hi-lighted in red.
\par\noindent

%% REVIEWER 1 and REVIEWER 2 COMMENTS HAVE TO BE SEPARATE.
Changes made in response to comments of Reviewer 2: 
\begin{enumerate}
%%1
\item
\textsf
{Page 3: The CLSVOF method is not an interface tracking one (in such methods
the interface is calculated simultaneously with the velocity and pressure
fields), but an interface capturing method (in such methods the interface is
reconstructed after the velocity field is calculated at each time step).
Although this is a matter of nomenclature, using the wrong name can be
confusing because the interface tracking algorithms, such as ALE, are
completely different from CLSVOF.}
\vspace{3 mm}

Response: \\
We agree with the reviewer. We have changed the ``interface tracking''
wording to ``interface capturing method''.
\\
Please see the 1st line from the top in ``I\hspace{-1.2pt}I\hspace{-1.2pt}I. A. Numerical method and governing equations''.
\\

%%2
\par\noindent
\item
\textsf
{Page 3: The definitions of the dimensionless numbers, Re, and Ca, are not
appropriate.  These numbers are defined here using the properties of air (the
dispersed material), whereas in the related references, which are used for
validation, they are defined using the properties of the continuous phase.
This is not so important in the validation tests where viscosity and density
ratios between the two phases are set equal to one, but it is important for the
new results. In general, the deformation of the inclusion is caused by the
forces exerted on it by the fluid in the matrix, and even more so when the
inclusion is just an air bubble, which has negligible density and viscosity.
Moreover, the capillary number is the ratio of viscous forces over capillarity
and the Reynolds number is the ratio of inertia over viscous forces, and the
viscous forces are negligible for air in comparison to viscous forces in the
liquid.}
\vspace{3 mm}

Response: \\
Thank you very much for catching this.  We have modified the definitions of
$Re$ and $Ca$ accordingly; it was an accident that we had previously 
defined these quantities in terms of the dispersed material.
The computations presented were performed using $Re$ and $Ca$ defined by the
density and the viscosity of the surrounding liquid.
\\
Please see Eq.(2).
\\

%%3
\par\noindent
\item
\textsf
{Page 3: The direction of gravity is not specified. Is it acting along the
z-direction? If so, see point \#9 below.  If not, in what direction does it
act? In other studies, an air bubble was considered under the action of both
gravity and velocity gradient in the same direction, ref. [3] and [4].
Interesting variations arise from the occurrence of a lift force on a
deformable bubble. Again, these additional works need to be cited and
discussed.}
\vspace{3 mm}

Response: \\
We have added the direction of gravity in Fig. 1 ($z$ direction);
albeit, we prescribe zero gravity in our simulations.  The rationale
for setting gravity equal to zero is given in more detail in our response
to item 9.  The purpose of our study is to do controlled
numerical experiments on the effects of the density and 
viscosity ratios.  We can isolate
the dependencies on these two ratios by prescribing zero gravity.  Although
[3] and [4] do not address the transition to break-up, they address a very
much related aspect to bubble motion in shear flow, so now we cite them in
the introduction.
\\

%%4
\par\noindent
\item
\textsf
{Page 4, top: “The singular Heaviside gradient term in the right hand side of
equation (5) is a body force...”. Although this term arises along with gravity
as a body force, in fact, it presumably acts only on the fluid/fluid
interface.}
\vspace{3 mm}

Response: \\
Following to reviewer's comment, we have provided an additional explanation
``The surface tension force expressed by the singular Heaviside gradient term
acts only on the gas-liquid interface.''
\\
Please see the 5th line from the top of Page 4.
\\

%%5
\par\noindent
\item
\textsf
{Page 4: Eq. (7a) determines the evolution of F, but the need to introduce F in
addition to the colour function required by VOF (which is determined by eq.
(7b)) is not discussed. F was not even mentioned before eq. (7). The authors
should also reference previous works where the need for this distinction has
been explained.}
\vspace{3 mm}

Response: \\
We agree with reviewer's comment. We have added details of the CLSVOF method at
the beginning of Section ``A. Numerical method and governing equations.''
\\

%%6
\par\noindent
\item
\textsf
{Page 5: ``The results shown in Figure 2 verify that our numerical approach can
reproduce the same drop breakup behavior presented in [8].'' The comparison is
only qualitative, not quantitative. One can readily see several differences in
all three cases presented.  The comparison could be improved either by
superimposing the plots or even better by extracting numerical values for the
velocity field or the drop shape. The latter may be more involved, which makes
addressing issue \#7 below even more important.}
\vspace{3 mm}

Response: \\
We agree that there are several differences between the previous
study by Renardy and Cristini [8] and our present study.  We claim
that the major cause of the differences is that Renardy and
Cristini [8] used a grid size of 
$\Delta x=R/8$ 
whereas in our case,
the effective fine grid resolution had a grid size of 
$\Delta x=R/16$.  We have made this more clear in the Figure 2 caption. 
\par\noindent
We have made {\em quantitative} comparisons for the Taylor deformation
parameter $D$ in Table I (drop deformation).  In the revision, we have
added another table with {\em quantitative} comparisons.  Please
refer to our newly added Table II 
(bubble deformation).  Note that it took 8 months to compute the results
reported in our new Table II.  
Comparisons of our results
against past studies 
(summarized in Tables I and II)
support the
validity of our bubble (and drop) deformation computations.
\\

%%7
\par\noindent
\item
\textsf
Page 6: It is well-known that interface capturing methods have drawbacks in
terms of accuracy, which in this problem could lead to uncertainty or
inaccuracy concerning the breakup conditions and the related bubble shapes.
Indeed, the topological changes may affect the small characteristic size of the
filament formed by the distorted drop or bubble, when it becomes comparable
with the grid size. For instance, in their study of two bubbles rising in line
(using VOF and Basilisk as a solver), Zang and Magnaudet, ref. [5], found that
refinement up to R/272 (i.e. the local grid size is 272 times smaller than the
bubble radius) is required in the proximity of the interface to properly
capture the topological changes of millimetric bubbles, since only then the
characteristic size of the grid is smaller than the average film thickness in
typical coalescence conditions. Instead, the authors use R/24 local grid size
at the most. Moreover, the Adaptive Mesh Refinement (AMR) technique is usually
employed for 3-4 consecutive levels of refinement, at least. In the present
work, only two levels of refinement have been used, why? The mesh convergence
study seems quite incomplete.  A more complete study with figures to verify it
should be included in an appendix concerning the mesh and time step
independence of the results. The type of time discretization used is not
stated; is it explicit and which algorithm is used?
\vspace{3 mm}

Response: \\
\par\noindent
In response to the question ``In the present
work, only two levels of refinement have been used, why?'', we added the 
following text to our revision, (under the heading ``selecting the 
appropriate grid size'')
\par\noindent

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.   
The accuracy of the ``Critical Reynolds Number''
depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures 8 and 11).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(i.e. ``Critical Reynolds number'')
(see Figures 3 and 12) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the ``critical Reynolds number''
accuracy.  
A smaller tolerance
would necessitate a larger supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

We make the distinction between our present research, and the research found
in the work of Zang, Ni, 
and Magnaudet[47,48] on predicting the
conditions for bubble mergers.  Even in the most extreme cases for
mergers, the largest Deformation parameter never exceeds
$0.4$ in [47].  In summary, our gridding requirements 
necessitate grid points distributed relatively evenly throughout the 
computational domain when a bubble is stretched to a $0.9$ Deformation,
whereas in [47] the gridding strategy necessitates 
a more localized strategy.

}

\par\noindent
In response to the enquiry ``The mesh convergence
study seems quite incomplete.  A more complete study with figures to verify it
should be included in an appendix concerning the mesh and time step
independence of the results.'', we have modified Figure 3 to report results
for three different grid resolutions.  Figure 3 illustrates that we
capture the proper break-up time as the grid is refined, and at the point
where the critical Reynolds' number is prescribed.  In Figure 3,
we point out that the time step was $O(10^{-5}$)s for the 
elongating bubble! ($\Delta x=R/16$) Further refinements reduce the
time step according to the surface tension time step constraint
(see section III(A)).  We now report in the revision the temporal
discretization of our method along with the timestep constraint.
Since surface tension is discretized explicitly in time, the
surface tension time step constraint is much more restrictive than
the classical CFL time step constraint of $\Delta x/U$.
(again see section IIIA).

%%8
\par\noindent
\item
\textsf
{\color{red} Several experimental studies concerning bubbles in shear flow
exist (for example, ref. [6]) so alternatively one could compare the numerical
results with experiments and avoid the comparison with other numerical results,
which are older and may not have achieved convergence due to lack of access to
the software and hardware we have today.}
\vspace{3 mm}

Response: \\
{\color{red} 
%As a quantitative comparison, we have newly made a comparison for bubble 
%deformation as Table I\hspace{-1.2pt}I:
We added a new quantitative comparison for bubble deformation in Table
I\hspace{-1.2pt}I to verify the validity of our computational results; in the
table we compare against the experimental results of Muller-Fischer et al.
(Experiments in fluids, 2008) and Rust and Manga (Journal of Colloid and
Interface Science, 2002). The new numerical results for bubble deformation
indicate that our computations are consistent with past studies.
%We believe that the effectivity of our computations could be proved for bubble deformation.
}
\\

%%9
\par\noindent
\item
\textsf
{The following is the most disturbing issue with this study: It is fine to
examine conditions leading to bubble breakup or other flow instabilities, but
the range of parameter values investigated must correspond to existing
materials and achievable flow conditions.  The authors have given the governing
equations in dimensional form and defined dimensionless numbers incorrectly as
stated in point \#2, but also assigned them values that may not be physically
relevant. Moreover, the bubble size is nowhere stated.  \\
Finally, the Reynolds number of the channel in this case would be: $Re_{\rm
channel} \simeq 6000$.  This high value of Re corresponds to turbulence
conditions.}
\vspace{3 mm}

Response: \\
As we have stated in response to your comment \#2, correct dimensionless
numbers based on the density and the viscosity of the surrounding liquid were
used in real computations.  Regarding physical properties and conditions in our
computations, we will give you an example for the case of $Ca$ = 0.3 and $Re$
=93.\\
\par\noindent
Physical properties 
(a silicone oil has a very similar physical properties with these values):\\
\quad $\rho_{\rm m}$ (density of liquid) = 1000 kg/m$^3$, 
$\mu_{\rm m}$ (viscosity of liquid) = 0.2 Pa$\cdot$s,
$\sigma$ (surface tension) = 0.025 N/m, $R$ (bubble radius) = 5.0 mm.\\
From ${\it \Gamma}$ = $2V/H$ = $2V/6R$, \  
$V$ = 3${\it \Gamma} R$ = $\frac{3\sigma Ca}{\mu_{\rm m}} $ = 1.1 m/s.
Also, we obtain $Fr$ (Froude number) = 
$\frac{{\it \Gamma}R}{\sqrt {gR}}$ = 1.7. \\
In this flow system, the effect of gravity may not be completely 
negligible. However, in our computations, 
we set $g = 0$ because we wanted to clearly isolate only the effects of 
the density and viscosity ratios. 
From another perspective, it can be said that we investigated 
phenomena in space.
At the end of the Problem Description section, 
we added a detailed explanation about physical properties and conditions.

Also, regarding the channel $Re$ number, we obtain $Re_{\rm channel} =
\frac{\rho_{\rm m}2VH}{\mu_{\rm m}}$ = 3300 for a simple linear shear-flow.  As
expected, $Re_{\rm channel} = 3300$ falls within turbulence conditions for
simple linear shear-flow of a single phase fluid.  It is important to note that
the generation of turbulence is suppressed and turbulent eddy dissipation
becomes large if bubbles exist in fluid flows. However, as far as we know,
there are no conclusive results on turbulence in a simple linear shear-flow
which includes bubbles.  We predict that the generation of turbulence is
considerably suppressed when bubbles occupy a significant portion of the
flow in a channel, as in this study. At present, we believe that our
computations based on the assumption of laminar flows are reasonable and not
unrealistic.\\

%%10
\par\noindent
\item
\textsf
{Page 9, Fig. 5 and related discussion: It is quite strange that the bubble
does not reach a final steady shape, but seems to expand, contract and then
expand again. Here, it is even more important to verify that the simulations
have converged (via mesh and time refinement) and to extend them much further
in time to determine if this periodic motion will prevail or finally lead to a
steady state.}
\vspace{3 mm}

Response: \\
{\color{red}  
We have made a new figure (Figure 8: shear stress profile) for
the cases of 
(i) $Re$ = 50 and $Ca$ = 0.3, and
(ii) $Re$ = 93 and $Ca$ = 0.3.
Our new Figure 8 is a part of our new 
Section~I\hspace{-1.2pt}V.C.  The purpose of this new material
is to gain further insight on the unsteady dynamics of bubble
elongation.  What we discovered was that for the Reynolds
number sufficiently below the critical value, a relatively quick
unsteady elongation period gives way to a steady state (with no
break up).  On
the otherhand for Reynolds number {\em close} to the critical Reynolds
number, there is a prolonged, unsteady, elongation period, in which
periodic motion is observed and the deformation parameter $D$ is close 
to one.  It is agreed, that the ``vacillating''
behavior cannot last forever, ultimately (perhaps stochastically!),
the bubble will either settle down or break.  We assert that regardless
of the outcome, this vacillating behaviour will always occur
in close proximity to the critical Reynolds' number.  In otherwords,
regardless of the outcome, we claim that one is assured of being within 
3 percent of the critical Reynolds number (see Figure 3).  In fact,
we hypothesize that there will {\em always} be ``vacillating'' behavior
if one is sufficiently close to the critical Reynolds number.  i.e.
given an almost infinite supply of computational resources, as one hones
in closer and closer to the critical Reynolds number, a ``tug of war''
will be observed between the surface tension force trying to pull the
bubble together versus the wall driven shear stress trying to pull the
bubble apart.  This discussion
is added to the paper.
}
%%11
\par\noindent
\item
\textsf
{ Page 13: Fig. 11, which distinguishes the two areas of bubble breakup against
just bubble deformation, presents only 4 points and a curve connecting them,
although this is an important physical result. Several more points must be
included in it to reach a definitive conclusion.}
\vspace{3 mm}

Response: \\
We appreciate the reviewer's opinion, but it is important to keep in mind that
at the start of our study there were no guiding initial results that indicated
even potential values for critical Reynolds numbers for a bubble. The critical
Reynolds numbers found and presented were obtained after extensive and
expensive computations.  After lengthy computations, we were able to determine
the bubble $Re_{\rm c}$ number for 4 $Ca$ numbers.  We believe
that a bubble breakup critical curve drawn based on 4 points will be 
useful for low $Ca$ numbers.  Following the suggestion from the 
other reviewer (reviewer 1), 
we have added
a drop breakup critical curve ($\lambda$ = $\eta$ =1.0) 
to Fig. 12 (formerly Fig. 11).  
The difference in $Re_{\rm c}$ number between the bubble and 
the drop is clearer in
the new Fig. 12.  
In future work, we will determine the $Re_{\rm c}$ for a 
variety of high $Ca$ numbers 
in order to construct a detailed critical curve 
for a wider range of $Ca$ numbers.\\

%%12
\par\noindent
\item
\textsf
{\color{red}  End of conclusions section: .........
These phrases state the obvious, the densities and viscosities are different between drops and bubbles and, therefore, they cause different dynamics. 
Instead, it is imperative that the authors give the different physics that the property differences generate. 
To this end, they should examine thoroughly and in every detail the velocity and stress fields in each arrangement. 
This may require additional figures and text, but it must be done. Without this examination and its conclusions, this study is more appropriate for a 
Journal on numerical methods in fluid flows than a physical review Journal. This is the second most important issue with this presentation}
\vspace{3 mm}

Response: \\
{\color{red}
We agree with reviewer's comment. In response, we have added 
two new images of the
shear stress profile for a bubble for two $Re$ conditions in Fig. 8.
A discussion of the new results, in terms of the 
shear stresses acting on the
bubble at the two different conditions, 
is given in Section~I\hspace{-1.2pt}V.C. 
}
\\

\end{enumerate}

Less important issues
\begin{enumerate}

\item
\textsf
{Page 2: The symbols for the domain length, width and height are L(x), W(y) and H(z), the variables in parenthesis indicate the direction related 
to the lengths L, W, and H. However, these are geometric parameters, and could be confused with functions, in the way they are given by the 
authors here and elsewhere in the text.\\
}
\vspace{5 mm}

Response: \\
We agree with reviewer's comment. We deleted $(x), (y), (z)$ and will simply
express the domain length, width and height as $L, W, H$.\\

\par\noindent
\item
\textsf
{Page 6: “... depicted in Figure 4(a), uses Ca = 0:3 and Re = 1:1.” It should be Figure 4(b).\\
}
\vspace{5 mm}

Response: \\
We've fixed the error.\\

\par\noindent
\item
\textsf
{Page 8: “... process is the almost same...” should be “... process is almost the same...”.\\
}
\vspace{5 mm}

Response: \\
We've fixed the typo.\\

\par\noindent
\item
\textsf
{The authors could cite another relevant paper where a similar analysis was undertaken [10]. 
The main difference is that the present computations are fully 3D, while in this suggested study shear 
flow until breakup is examined of a 2D bubble. Nevertheless, there are some similarities 
that should be recognized. The present work is more accurate and extensive.\\
}
\vspace{5 mm}

Response: \\
The paper by Wei et al. is now cited in the Introduction.\\

\par\noindent
\item
\textsf
{The authors should use D instead of De as the symbol for the parameter introduced by Taylor to 
determine the extent for bubble deformation, because De stands for the Deborah number.\\
}
\vspace{5 mm}

Response: \\
We have changed the symbol from $De$ to $D$ as suggested.\\

\end{enumerate}

\closing{On behalf of the authors,}
Mitsuhiro Ohta

\end{letter}
\end{document}