### 2024-11-04 Revised-paper/Reply/Reply-Reviewer1.tex
significance and novelty

%  -----------------------------------------------------------------------------
%  Modified:October 2024
%  -----------------------------------------------------------------------------
\documentclass{letter}

 \date{\today}

\usepackage{color}

\begin{document}
\begin{letter}{
Chemical Engineering Science\\
Professor Wei Ge, Editor\\}

\opening{Dear Prof. Ge,}

Please find enclosed the revised version of our manuscript (CES-D-24-01867):
``Numerical Simulation of Bubble Deformation and Breakup under 
Simple Linear Shear Flows''.
We thank the reviewers for their comments and thoughts regarding improvement 
of our paper. We believe that we have addressed all of the reviewers’ 
concerns; the changes are itemized in detail below.

\par\noindent
NOTE: all references to equations and figures below are with
reference to the numbering scheme of the revised version of the paper,
not the original version.  Also, changes are hi-lighted in red.
\par\noindent

Changes made in response to comments of Reviewer 1: \begin{enumerate}
%%1
\item
\textsf
{Abstract, hightlights, and introduction are written in a way that confuses the reader: the text jumps back and forth between bubbles and drops as well as shear flows and other flow configurations. Then it suddenly moves to lift and deformation rather than breakup. Finally a number of random applications are mentioned without any relation to bubble breakup. The authors should clearly state what their contrbution is from the outset. The reader will not go on until the end to figure that out by himself. The use of shorter paragraphs each addressing a single issue is very helpful.}
\vspace{3 mm}

Response: \\
In the revision, we made many improvements to the grammar and structure so that the reader can follow in a much more efficient manner. \\
\\

%%2
\item
\textsf
{Method: The use of only 2 levels of grid refinement appears odd. For an accurate representation of the flow near the interface a very fine mesh is typically needed. Only 32 cells over the bubble diameter is hardly sufficient ot get accurate results as corroborated by the mediorce agreement of the presented simulation results with the experimental data used for validation. Two more levels of grid refinement are likely needed to resolved the flow near the bubble while farther away some coarsening may be possible. Optimizing the grid is imperative in a situaltion like the considerd one due to the strong deformation of the bubble which high curvature of the interface and the singular nature of the breakup event. Fully capturing the breakup, which represents a topological singularity, is impossible without a subgrid scale model.}
\vspace{3 mm}

Response: \\
The grid size and adaptive meshing strategy we adopted are chosen to answer the research question as to the conditions that determine 
whether a bubble in shear flow will break up.  
In such a case, we must accurately capture the balance of forces concerning the (non-local) force exerted from the wall-driven 
flow acting against the interfacial surface tension force.

The accuracy of the ``Critical Reynolds Number'' depends on the largest Taylor Deformation parameter $D$ supported by the grid.  
As we report here, we have found that as long as the grid size is fine enough to support a Taylor Deformation parameter $D<0.95$, 
then the transition region (i.e., ``Critical Reynolds number'') (see Figures 3 and 12) will be captured with a tolerance of three percent.

Figure 3 illustrates that we capture the proper break-up time as the grid is refined and at the point where the critical Reynolds
number is prescribed. In Figure 3, we point out that the time step was $O(10^{-5}$)s for the elongating bubble ($\Delta x=R/16$).
Further refinements reduce the time step according to the surface tension time step constraint. 
The simulation time would become impractical if we tried to improve the accuracy of the ``critical Reynolds number'' further.  
(For example, in the present study, the computation took more than one year for an experiment with $Re$ = 0.0 and $Ca$ = 1.63 
(M\"{u}ller-Fischer et al. (2008)).
A smaller tolerance would necessitate a larger  supported Deformation parameter $D$, which would, in turn, necessitate a 
higher aspect ratio computational domain, increased droplet surface area at break-up, increased number of time steps, and 
higher resolution for representing the drop/bubble at its thinnest point. \\
\\

%%3
\par\noindent
\item
\textsf
{Results: Mostly just pictures are shown from which little is to be learned. The only results of which future use can be expected is the regime map in Fig. 12. Bu this is also only qualitative and based on very few data points. More combinations of Ca and Re should be provided and a quantitative fit to the regime boundary be obtained. Moreover only two values of density and viscosity ratio equal to 1 and ~0.001 are used. An investigation of intermediate values would be straight forward and would allow to provide also models for the influence of these obviously important parameters. In particular it should be clarified whether the difference in behavior is more due to the density ration or more due to the viscosity ratio. Also it needs to be investigated how small these ratios need to be for the gas phase properties to be considered as negligibly small. For the practically important case of air bubbles in water the viscosity ratio is 0.02, an order of magnitude larger than considered in the paper. Whether this can already be considered negligible or not remains unclear.}
\vspace{3 mm}

Response: \\
		We cannot understand this criticism in the context of the referee's previous comment.  They are contradictory statements.  In the previous comment, the referee complains that we have lack of results, and in this section the referee complains that our results are superfluous.  Every Figure that we show strategically informs the reader about the tensile strength properties of a bubble in wall driven shear flow.  All the results that we report are novel and significant tion, and in the present comment, the referee complains about our results which clearly highlight why this problem is so hard, and why it is unreasonable to ask for more.  Already the presented results that we present regarding the ``tensile strength'' of a {\em bubble} in the presence of driven walls
We believe the figures and computational results presented in this study are very valuable to developing the study of drop/bubble deformation and breakup.
The following points were especially presented and revealed in this paper for the first time. \\
\begin{enumerate}
\item The behavior of the bubble deformation process around critical $Re$ conditions with $Ca$ = 0.3  $\sim$ 1.0 has been presented for the first time.
\item The behavior of the bubble breakup process at critical $Re$ conditions with $Ca$ = 0.3  $\sim$ 1.0 has been presented for the first time.
\item The value of critical $Re$ numbers was found for $Ca$ = 0.3  $\sim$ 1.0 for the first time.
\item  We showed that most previous computational studies overestimated drop deformation and breakup because most previous studies used a domain size with $W$ (width ($y$directional length)) =4$R$. Drop deformation is promoted when using a domain size with $W=4R$. 
\item The difference in the deformation and breakup between the bubble and the drop with $\lambda = 1$ and $\eta = 1$ was clearly shown.
\item Shear-induced bubble/drop breakup critical curves were obtained. 
\end{enumerate}

Regarding the effect of $\lambda$ and $\eta$, ongoingly we have examining the effect of $\lambda$ and $\eta$ on drop breakup for following 
combinations of $\lambda$ and $\eta$:

\begin{tabbing}
 \hspace{55mm} \= \hspace{10mm} \kill
 \hspace{5mm} 1. $\lambda$ = 1.0, $\eta$ =1.0 (done) \> 2. $\lambda \simeq 0.0$, $\eta \simeq 0.0$ (this study) \\ 
 \hspace{5mm} 3. $\lambda$ = 1.0, $\eta \simeq 0.0$ (done) \> 4. $\lambda$ = 1.0, $\eta$ =0.01 \\
 \hspace{5mm} 5. $\lambda$ = 1.0, $\eta \simeq 0.1$ \> 6. $\lambda$ = 0.1, $\eta \simeq 0.0$ (done) \\
 \hspace{5mm} 7. $\lambda$ = 0.1, $\eta$ =0.01 \> 8. $\lambda = 0.1$, $\eta$ =0.1  \\
 \hspace{5mm} 9. $\lambda$ = 0.1, $\eta$ =1.0 (done) \> 10. $\lambda$ = 1.0, $\eta$ =10  \\
 \hspace{5mm} 11. $\lambda$ = 1.0, $\eta$ =100 \> 12. $\lambda$ = 1.0, $\eta$ =1000 \\
\end{tabbing}

Since it takes a very long computational time to obtain results for the effect of $\lambda$ and $\eta$ on drop deformation and breakup, 
all computations for these conditions have not been finished. We will also present the fruits of research on the effect of density and 
viscosity ratios in the next papers. \\
\\

%%4
\par\noindent
\item
\textsf
{The effect of gravity is neglected which potentially obliterates the results. It is not at all clear that the effect of the shear flow can be separated from that of gravity without rendering the results meaningless for any practical application. Even when one wants to do so it needs to be considered that the bubbles will be deformed into an ellipsoidal shape. Then at least the influence of initial condition has to be checked in order to ensure that the results make any sense. \\}
\vspace{3 mm}

Response: \\
In our computations, we set $g = 0$ because we wanted to clearly isolate only the effects of the density and viscosity ratios and 
wanted to compare previous studies dealing with the deformation and breakup of a drop with $\lambda =1$ and $\eta = 1$.
When the gravity is considered, we obtain $Fr$ (Froude number) $\left(= \frac{{\it \Gamma}R}{\sqrt {gR}} \right)$ = 1.7 ($Re$ = 93 and $Ca$ = 0.3) 
and 1.9 ($Re = 43$ and $Ca$ = 0.8). \\
Although the values of $Fr$ are not so large, the effect of gravity (bubble rise motion) may not be completely negligible in terms of $Fr$.
However, bubbles for both conditions reach the breakup at $t$ = about 0.5 s. 
($V$ (moving wall speed) had about 1.1 $\sim$ 1.3 m/s for the range of $Ca$ and $Re$ in our computaions.)
Therefore, the effect of gravity (bubble rise motion) is expected to be negligible to the behavior of bubble deformation and breakup 
around critical $Re$ number conditions. Our computational results will provide fundamental knowledge and findings for practical applications.
In the revised paper, we have described these explanations.\\
\\
Please see lines 123 to 130.
\\

%%5
\par\noindent
\item
\textsf
{Language and writing (examples):\\
line 191 and following: "Timestep $\Delta t$" appears out of any context. So do "Step 1." to "Step 4".\\
line 232: ""bubble deformation in simple linear shear flow" results" is impossible language. \\}
\vspace{3 mm}

Response: \\
We carefully checked the paper and modified poor English writing, expressions, etc.\\
\textcolor{red}
{
Mark and Edwin: Please modify these sentences at line 191 and line 232. (line198 and 242 in the revised paper.)\\
}
\\
Please confirm lines  198 to 
\\

\end{enumerate}

\closing{On behalf of the authors,}
Mitsuhiro Ohta

\end{letter}
\end{document}