### 2024-11-04 Revised-paper/Highlights.tex

%  -----------------------------------------------------------------------------
%  Modified:October 2024
%  -----------------------------------------------------------------------------
\documentclass{article}

\usepackage{color}

\begin{document}
Highlights\\

\begin{itemize}
\item The deformation and breakup of a bubble in liquids undergoing simple linear shear flow are computationally reproduced.

\item Shear-induced bubble deformation and breakup process differs significantly from that of the analogous drop system under similar flow conditions.

\item The critical conditions for bubble breakup concerning the Reynolds and capillary numbers are determined.

\item The bubble oscillates between elongating and shrinking motions for non-rupturing flow conditions.

\end{itemize}

\end{document}